A Closer Look at Time-Preserving Walks on Evolving Graphs

### 2018-04-11 paper.tex
make the idea clear

\title{A Closer Look at Time-Preserving Paths on Evolving Graphs}

@@

\begin{abstract}
This paper studies evolving graph centrality from the perspective of time-preserving paths. We describe time-preserving paths based on two dimensions of variation, namely static edges and causal edges.

A time-preserving path on evolving graphs makes implicit assumptions about the type of edges that a path traverses. For example, the path can count only static edges or can take account of both static edges and causal edges. In addition, the length of a time-preserving path can be either spatial or temporal.

We study the impact of different time-preserving paths on node centrality by varying the edge weights of static and causal edges.
We also compare evolving graph centrality with the aggregated static graph case to gain insight about the advantage of an evolving graph model.
We construct an evolving coauthor network from a collection of research papers for the experiments.
\end{abstract}

@@

In the 1985 science-fiction film ``Back to the Future'', Marty McFly (played by Michael J. Fox) travelled back in time and accidentally changed history.
However, we cannot travel back in time or even send a message to the past.
An evolving graph in general is a time-dependent graph $G(t) = (V(t), E(t))$.
To traverse an evolving graph, one has to consider time. In particular, a path on an evolving graph needs to be time-preserving, meaning a node can not pass a massage (through an edge) to a node at a earlier time.

% time-respecting path
We represent an evolving graph as a time-ordered sequences of graphs, similar to the work of Tang and coworkers \cite{nicosia13, tang09, tang102, tang10} and Grindrod, Higham and coworkers \cite{grindrod13,grindrod11}. The idea is to divide the system's time span into temporal slices and regards it as a sequence of static graphs $G^{[t]}$, one for each layer. A node on a slice $G^{[t]}$ is represented by a pair $(v, t)$, where $v$ is a node of
$G^{[t]}$ and $t$ is the time stamp of the slice. A time-respecting path of length $n$ is a sequence of these node pairs $\langle (v_1, t_1), (v_2, t_2), \ldots ,(v_m, t_m)\rangle$, where $t_1 \le t_2 \le \cdots \le t_n$. Note if $t_i = t_j$, then $\langle (v_i, t_i), (v_j, t_j) \rangle$ is an edge at a temporal slice,  otherwise it is an edge linking nodes from different temporal slices.
We call edges of the first kind \emph{static edges} and of the second kind
\emph{causal edges}.
For example, in Figure \ref{fig:eg_shortest_path}
$\langle (1, t_1), (2, t_1) \rangle$ is a static edge at time stamp $t_1$ and $\langle (3, t_2), (3, t_3) \rangle$ is a causal edge between time stamp $t_2$ and time stamp $t_3$.
The difference between static edges and causal edges were first explicitly considered in \cite{chen16}.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
  \node (w1) at (-6, 9) {$w_1$};
  \node (w2) at (3, 13) {$w_2$};
  \node (w3) at (-2, 8) {$w_3$};
  \node (w4) at (2, 4) {$w_4$};
  \node (w5) at (8.5, 7) {$w_5$};s
  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
   \draw[every edge,->] (\from) -- (\to);
 \draw[dotted,->](n4) -- (n12);
\draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows and directed edges between graphs are shown as dotted arrows, where $w_1, w_2 \ldots w_5$ are edge weights.}
\label{fig:eg_shortest_path}
\end{figure}

It is natural to define the length of a path on a static graph as the number of nodes travelled through the path. For the weighted graph case, the length of a path is the total weight of all the edges in the path. We observe that static edges link nodes in space while causal edges link nodes in time. Hence we would like to distinguish temporal distance from spatial distance on an evolving graph: the length of a static edge is equal to the edge weight (or $1$ in the unweighted graph case); the length of a causal edge is the time distance between the two nodes. In the simplest case, the time-respecting path
$\langle (1, t_1) ,(2, t_1) , (2, t_3), (3, t_3)\rangle$ in Figure \ref{fig:eg_shortest_path} has length $4$ in total because two static edges $\langle (1, t_1) (2, t_1) \rangle$ and $\langle (2, t_3), (3, t_3) \rangle$ have length one, whereas $\langle (2, t_1) (2, t_3) \rangle$ traverse $2$ time stamps so has length $2$.

% We studies evolving graph centrality from the perspective of temporal network flows.
% Researchers have identified the time-preserving nature of such network flows, i.e.,
% we cannot travel between nodes from a later time to an early time.
% However, a closer look tells us there are subtle differences
% in previous papers

% How about a reverse time-preserving network flow?

% our contribution this paper
Borgatti noted in \cite{borgatti05} that the manner of traffic flow entails a new way to think about centrality. In particularly, the importance of a node in a network can not be determined without reference to how traffic flows through the network.
To develop a centrality algorithm on evolving graphs, one has to consider the manner of temporal traffic flow.
A time-preserving path on evolving graphs can be described by static edges, causal edges, or both.
Grindrod et al.'s definition of dynamic walk \cite{grindrod11} only counts static edges.
Tang et al. \cite{tang10s} measure a temporal shortest path in time unit (or causal edges) only. If a node can be reached by another node in $k$ time stamps then the distance between the two nodes is $k$. Chen and Zhang \cite{chen16} considered both static edges and causal edges in their definition of temporal path.

The aims of this paper are as follows. First, to identify the type of time-preserving path in a list of common evolving graph centrality measures. Second, to vary static and causal edge weights and analyze the impact on selected measures of centrality.
Third, to solve a real world problem using evolving graph centrality: identify key authors in an evolving coauthor network.
We carry out the experiment using the Julia dynamic network analysis package
EvolvingGraphs.jl (\url{https://github.com/EtymoIO/EvolvingGraphs.jl}).
We use the research paper data gathered from Etymo (\url{https://etymo.io}), a visual search engine for data scientists.

% \section{Background}
% \label{sec:preliminaries}
%
% Let $G$ be a directed graph with $n$ nodes with the corresponding adjacency matrix
% $A \in \R$.
%
% A \textbf{walk} of length $r$ from node $i_1$ to node $i_{r+1}$ is a sequence of edges
% $(i_1, i_2), (i_2, i_3), \ldots (i_r, i_{r+1})$.
%
%
% \subsection{Graph Centrality}
% \label{sec:katz-centrality}
%
% % definition of various graph centrality algorithms
% We review the Katz centrality here.
% Let $A$ be the adjacency matrix of a graph $G=(V,E)$ and $\rho(A)$ be the spectral radius of $A$. If $\alpha < 1/ \rho(A)$, then the Katz centrality vector of $G$ is
% \[
% (I - \alpha A)^{-1}e.
% \]
%
% The \emph{betweenness centrality} of a node $i$ in a static graph is defined as
% $$
% C_i = \sum_{j \in V}\sum_{k \in V, k \ne j} \frac{\rho_{jk}(i)}{\rho_{jk}},
% $$
% where $\rho_{jk}$ is the number of shortest paths from node $j$ to node $k$, while $\rho_{jk}(i)$ is the number of shortest paths from node $j$ to node $k$ that pass through node $i$

% The \emph{communicability betweenness centraltiy} \cite{estrada09} for a node $i$ is defined as
% $$
% C_i = C_N\sum_p \sum_q \frac{\exp(A)_{p,q} - \exp(A - E(i))_{p,q}}{\exp(A)_{p,q}},
% $$
% where $C_N = \frac{1}{(N-1)^2 - (N-1)}$ is a normalizing factor and $E(i)$
% has nonzeros only in row and column $i$, and in such row and column has $-1$ wherever $A$ has $1$. Similarly, the \emph{resolvent betweenness centrality} for node $i$ is defined as
% $$
% C_i = C_N\sum_p \sum_q \frac{(I - \alpha A)^{-1}_{p,q} - (I - \alpha(A - E(i)))^{-1}_{p,q}}{(I - \alpha A)^{-1}_{p,q}}.
% $$
% \subsection{Eigenvector Centrality}
% \label{sec:eigenv-centr}

% Given an adjacency matrix $A$ of a graph $G$, eigenvector centrality is defined as the dominant eigenvector of $A$, i.e.,
% \[
% \lambda v = Av,
% \]
% where $\lambda$ is the dominate eigenvalue of $A$ and $v$ is the corresponding eigenvector. Variations of eigenvector centrality includes Katz centrality \cite{katz53}, PageRank \cite{page99}, Hubbell's clique identification \cite{hubbell65}, Friedkin's actors' network centrality \cite{friedkin91}.
% In general, all eigenvector-like centrality are influence measures, meaning that even if a node $a$ only influences one node $b$, who influences many other nodes, then $a$ is highly influential.

% Let $A$ be the adjacency matrix of a graph $G = (V,E)$, i.e., $A_{i,j} = 1$ if node $i$ is linked to node $j$ and $A_{i,j} = 0$ otherwise. The idea of eigenvector centrality is that
% each node's centrality is the sum of the centrality values of the nodes that it is connected to. In other words, the relative centrality of node $v$ can be defined as
% \[
% x_v = \frac{1}{\lambda}\sum_{t \in M(v)} x_t = \frac{1}{\lambda}\sum_{t \in G} A_{v, t}x_t,
% \]
% where $M(v)$ is the set of neighbor nodes of $v$. We can write it in matrix notation as
% \[
% \lambda x = Ax.
% \]
% The eigenvector centrality of $G$ is the dominant (or principle) eigenvector of $A$.
% One may ask why we choose dominant eigenvector not any other eigenvectors? The answer comes from Perron-Frobenius Theorem (more details later).

% In \cite{bonacich91}, Bonacich prove the following theorem.
% \begin{theorem}
%   Assume that a symmetric non-negative matrix $A$ has a single largest eigenvalue $\lambda_1$ and its corresponding eigenvector $v_1$. Let $e$ be a column vector of ones. The infinite sum $A + \beta A^2 + \beta^2 A^3 + \beta^3 A^4 + \cdots$ converges
% if $|\beta| < 1/ \lambda_1$. Moreover, as $\beta$ approaches $1/\lambda_1$ from below,
% $(1-\beta\alpha_1)(A+ \beta A^2 + \beta^2 A^3 + \beta^3 A^4 + \cdots)e$ approaches $v_1$.
% \end{theorem}
% In other words, the eigenvector centrality of $G$ can be thought as proportional to an
% infinite sum
% \begin{equation}
% \label{eq:s}
% Se = Ae+ \beta A^2e + \beta^2 A^3e + \beta^3 A^4e + \cdots.
% \end{equation}
% Note however if $\beta = 1/\lambda_1$, the series does not converge.
% Eq. \eqref{eq:s} offers great flexibility to think about related centrality measures.
% For example, in the PageRank case
% \[
% r = Gr,
% \]
% where $G$ is a stochastic matrix. Then
% \[
% Se = Ge + \beta G^2e + \beta^2 G^3e + \beta^3 G^4 e + \cdots
% \]
% converges to $r$ as $\beta$ approaches $1$ from below.

% Note when $\alpha < 1/\rho(A)$, the spectral radius of $A$,
% \[
% (I - \alpha A)^{-1} = I + \alpha A + \alpha^2 A^2 + \alpha^3 A^3 + \cdots.
% \]
% Hence from eq. \eqref{eq:s} we have
% \[
% \beta Se + 1 = (I - \beta A)^{-1}e.
% \]
% This means eigenvector centrality can interpreted as Katz centrality with special chosen parameter value $\beta$. We mention that Aprahamian et. al. \cite{ahh16} consider selecting the Katz parameter based on the objective of matching the centrality of the exponential counterpart.

% \textbf{WZ:} some analysis need to be done on the relationship between
% $\beta$ and $v_1$. Derive some upper or lower bounds.

% %katz case
% Suppose we let
% \[
% S = A + \alpha_1^{-1} A^2 + \alpha_1^{-2} A^3 + \cdots,
% \]
% then $u = Se$ is proportional to the eigenvector $v_1$.

% In PageRank's case
% \[
% r = Gr,
% \]
% which means $\alpha_1 = 1$.

% Since Katz centrality can be written as
% \[
% (1- \alpha A)
% \]
% so it can be regard as a special case of eigenvector centrality.

% % \subsection{The Datasets}
% % \label{sec:datasets}

% % The data is a list of papers with associated information, including author names, affiliation, journal, and year of publication, and the full text content. We will test our algorithms on a collection of articles from, NIPS, an
% d ICLR.

% \subsection{Author Name Disambiguation}
% \label{sec:auth-name-disamb}

% We follow the idea from \cite{martin13} to disambiguate author names.

% \section{Eigenvector Centrality on Evolving Graphs}
% \label{sec:centr-algor}

% Let $A^{[t]}$ be the corresponding graph $G^{[t]}$ at a time label $t$. In \cite{grindrod11},
% Grindrod et. al. observes that the matrix product $A^{[t_1]}A^{[t_2]}\cdots A^{[t_w]}$ has
% $(i,j)$ element that counts the number of dynamic walks of length $w$ from node $i$ to node $j$ on which the $m$-th step of the walk takes place at time $t_m$. Therefore, the Katz centrality for evolving graphs can be written as
% \[
% Q = (I - \alpha A^{[0]})^{-1}(I - \alpha A^{[1]})^{-1} \cdots (I - \alpha A^{[m]})^{-1}.
% \]

% Follow the discussion from Section \ref{sec:eigenv-centr}, we can generalize eigenvector centrality and PageRank on evolving graphs.

\section{Time-Preserving Paths}
\label{sec:time-pres-paths}

We now describe time-preserving paths in more detail. We follow the definition of
\emph{evolving graph}, \emph{temporal node} and \emph{active node} in \cite{chen16}.

@@

Note the node sets $V^{[t]}$ can change over time, i.e., nodes may appear or disappear at a particular time stamp.
For example, in Figure \ref{fig:eg_shortest_path}, at time stamp $t_1$, $V^{[1]} = \langle 1, 2 \rangle$ and $E^{[1]} = \langle (1,2) \rangle$. Each graph $G^{[t]}$ can be represented by its adjacency matrix $A^{[t]}$.
We can represent $G_n$ by a list of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{[n]} \rangle$.

@@

In Figure \ref{fig:eg_shortest_path}, the filled green circles are active nodes. For the adjacency matrix representation, a node $i$ is active at time stamp $t$, then
the $i$th row or the $i$th column of $A^{[t]}$ has at least one none-zero entry.
 In general, a time-preserving path can be defined as follows.

\begin{definition}
A \textbf{time-preserving path}  of length $m$ on an evolving graph $G_n$ is a time-ordered sequence of active nodes, $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$, where $t_1 \le t_2 \le \cdots \le t_m$ and
$v_i = v_j$ if and only if $t_i \ne t_j$.
\end{definition}

@@

We can denote a \emph{weighted static edge} as $\langle (v_i, t), (v_j, t), w_s \rangle$ and a \emph{weighted causal edge} as $\langle (v, t_i), (v, t_j), w_t \rangle$.
Here $w_s$ represents the spatial distance between the two nodes $v_i$ and $v_j$; $w_t$ represents the
temporal distance of $v$ at different time stamps.
In Figure \ref{fig:eg_shortest_path}, the black arrows represent static edges and the dotted arrows represent causal edges.
Suppose we assign all the static edge weight to be $1$ and
assign the causal edge to be the number of time stamps between the pair. Then in Figure \ref{fig:eg_shortest_path}
the edge weight between $(1, t_1)$ and $(2, t_1)$ is $1$ and the edge weight between $(2, t_1)$ and $(2, t_3)$
 is $2$.

A time-preserving path can traverse both causal edges and static edges.
We will see in the next section that evolving centrality measures make implicit assumptions about the type of edges in the
time-preserving path that is used.

Recall a time-preserving path is a time-ordered sequence of active nodes. Here we define a \emph{reversed temporal path} to be a reversed time-ordered sequence of active nodes, i.e.,
$t_1 \ge t_2 \ge \cdots \ge t_m$.

\section{Relation to Centrality Measures}
\label{sec:topol-temp-flow}

In \cite{tang10s}, a temporal shortest path is measured in time units only, i.e., if node
$v_i$ can reach node $v_j$ in $k$ time steps then the distance between node
$v_i$ and $v_j$ is $k$. We can apply temporal breadth first
search (BFS) \cite{chen16} to discover the temporal paths and then only take account of the temporal difference between the first node $(v_1, t_1)$ and the last node $(v_n, t_n)$, i.e., $t_n - t_1$.

If we use this definition of temporal path for the temporal Katz centrality introduced in
\cite{grindrod11}, we have a new centrality measure.
In the matrix form, the nonzero entries of $A^{[t_i]} A^{[t_{i+1}]}$ counts all the temporal paths of length $1$, and
$A^{[t_i]}\cdots A^{[t_j]}$, where $i < j$ counts all the temporal paths of length $j -i$.
For example, $A^{[1]}A^{[4]}$, $A^{[1]}A^{[1]}A^{[4]}$ and $A^{[1]}A^{[2]}A^{[3]}A^{[4]}$ are all walks of length 3.
In the original Katz centrality, the `attenuation' factor $\alpha$ is imposed on the
number of walks, i.e., longer walks has smaller effect to the overall rating.
Hence the final result is the summation of all products of the form
\[
\alpha^k A^{[i]} \cdots A^{[i+k]}
\]
% Note this can not be written in \eqref{eq:katz}, as discussed in Section \ref{sec:temp-katz-centr}.

Another definition of temporal path was introduced by \cite{chen16}. Here
both static edges and causal edges are considered. We can represent an evolving graph by a block adjacency matrix, as shown in Section \ref{sec:centr-block-adjac}. Centrality on this block matrix gives the importance of each node at all time stamps. We can average these rating at different time stamps to
get the final rating.

We review a few well-known measures of evolving graph centrality.
(Design a simple example and discuss it in each case below.)
\subsection{Temporal Katz Centrality}
\label{sec:temp-katz-centr}

@@

\subsection{Temporal Betweenness Centrality}
\label{sec:temp-betw-centr}

Recall that betweenness centrality of a node in a static graph is defined by counting the fraction of shortest paths that traverse it.
We could extend
betweenness centrality on evolving graphs by replacing shortest paths with temporal shortest paths \cite{nicosia13}.
\begin{equation}
  \label{eq:bet1}
C_i^{betweenness} = \sum_{j \in V}\sum_{k \in V, k \ne j}\frac{\alpha_{jk}(i)}{\alpha_{jk}},
\end{equation}
where $\alpha_{jk}$ represents the number of temporal shortest paths from node $j$ to node $k$ and $\alpha_{jk}(i)$ represents the number of temporal shortest paths from node $j$ to node $k$ that pass through the node $i$.

>>>>

\title{A Closer Look at Time-Preserving Walks on Evolving Graphs}

@@

\begin{abstract}
This paper studies evolving graph centrality from the perspective of time-preserving walks. We describe time-preserving walks based on two dimensions of variation, namely static edges and causal edges.

A time-preserving walk on evolving graphs makes implicit assumptions about the type of edges that a walk traverses. For example, the walk can count only static edges or can take account of both static edges and causal edges. In addition, the length of a time-preserving walk can be either spatial or temporal.

We study the impact of different time-preserving walks on node centrality by varying the edge weights of static and causal edges.
We also compare evolving graph centrality with the aggregated static graph case to gain insight about the advantage of an evolving graph model.
We construct an evolving coauthor network from a collection of research papers for the experiments.
\end{abstract}

@@

In the 1985 science-fiction film ``Back to the Future'', Marty McFly (played by Michael J. Fox) travelled back in time and accidentally changed history.
However, we cannot travel back in time or even send a message to the past.
To traverse an evolving graph, a.k.a a time-dependent graph, one has to consider time. In particular, a walk on an evolving graph needs to be time-preserving, meaning a node can not pass a massage (through an edge) to a node at a earlier time.

% time-respecting walk
We represent an evolving graph as a time-ordered sequences of graphs, similar to the work of Tang and coworkers \cite{nicosia13, tang09, tang102, tang10} and Grindrod, Higham and coworkers \cite{grindrod13,grindrod11}. The idea is to divide the system's time span into temporal slices and regards it as a sequence of static graphs $G^{[t]}$, one for each layer. Such representation arises naturally in many real-world applications. For example,
consider networks of users interacting through messaging. Each interaction between two users $A$ and $B$ are stored with a time stamp $t$. If we represent a user as a node then a user $A$ at time stamp $t$ is represented by a pair $(v,t)$ and a message sent from user $A$ to user $B$ at time stamp $t$ can be represented as $\langle (A, t) (B, t) \rangle$.

In general, a node $v$ on $G^{[t]}$ is represented by a pair $(v, t)$, where $t$ is the time stamp of the temporal slice.
A sequence of interactions between users form a walk. Suppose $A$ sent a message to $B$ at time stamp $t_1$
and then $B$ sent this message to $C$ at time stamp $t_2$. We can represent this activity as $\langle (A,t_1), (B,t_1), (B, t_2), (C,t_2)\rangle$, where each pair of two nodes has meanings: $\langle (A,t_1), (B,t_1) \rangle$ and $\langle (B,t_2), (C,t_2) \rangle$ represent interactions between users $A$, $B$, and $C$ and $\langle (B,t_1), (B,t_2) \rangle$ represents the fact that user $B$ holds the message at time stamp $t_1$ and $t_2$.

A time-respecting walk is a sequence of $\langle (v_1, t_1), (v_2, t_2), \ldots ,(v_m, t_m)\rangle$, where $t_1 \le t_2 \le \cdots \le t_n$ are the corresponding time stamps. If $t_i = t_j$, then $\langle (v_i, t_i), (v_j, t_j) \rangle$ is an edge where both nodes are on the same temporal slice,  otherwise it is an edge linking nodes from different temporal slices.
We call edges of the first kind \emph{static edges} and of the second kind \emph{causal edges}.
Recall in static graph, a walk of length $m$ is defined as a sequence of nodes $v_1, v_2, \ldots, v_{m+1}$ such that for each $i = 1, \ldots, m$ there is an edge from $v_i$ to $v_{i+1}$. How to determine the length of a time-respecting walk?
One could define the length of a time-respecting walk as the total number of temporal slices (see shortest temporal distance in \cite{tang09}) or the number of static edges (see dynamic walk in \cite{grindrod11}). This paper studies the impact of these two different definitions. In particular, we studies a general definition of temporal distance that takes account of both temporal slices and static edges.

% Related works
Researchers have generalised many static graph centrality algorithms for evolving graphs. One approach is to use the fact that the elements of the matrix product of adjacency matrices at different time stamps can correctly count the number of temporal walks.
However, not every centrality algorithm can be generalised in this way. Another approach is to define evolving graph centrality algorithms using time-respecting walks. This approach is applicable to algorithms that are defined using paths or walks.

% no work on pagerank so far. By understanding temporal walk, we could provide a generalization of pagerank on evolving graphs.

In Higham's following up papers, time delay is considered as alternative format of Katz centrality.
But the impact of temporal is left unconsidered. We studied the problem on centrality from two different types of centrality: one that depends on matrix unfolding and one that depends on temporal walk.

% our contribution this paper
Inspired by work
Borgatti noted in \cite{borgatti05} that the manner of traffic flow entails a new way to think about centrality.
where the author simulate different traffic flows to study centrality.

In particularly, the importance of a node in a network can not be determined without reference to how traffic flows through the network.
To develop a centrality algorithm on evolving graphs, one has to consider the manner of temporal traffic flow.
A time-preserving walk on evolving graphs can be described by static edges, causal edges, or both.

% Grindrod et al.'s definition of dynamic walk \cite{grindrod11} only counts static edges.
% Tang et al. \cite{tang10s} measure a temporal shortest walk in time unit (or causal edges) only.
% If a node can be reached by another node in $k$ time stamps then the distance between the two nodes is $k$. Chen and Zhang \cite{chen16} considered both static edges and causal edges in their definition of temporal walk.

The aims of this paper are as follows. First, to identify the type of time-preserving walk in a list of common evolving graph centrality measures. Second, to vary static and causal edge weights and analyze the impact on selected measures of centrality.
Third, to solve a real world problem using evolving graph centrality: identify key authors in an evolving coauthor network.
We carry out the experiment using the Julia dynamic network analysis package
EvolvingGraphs.jl (\url{https://github.com/EtymoIO/EvolvingGraphs.jl}).
We use the research paper data gathered from Etymo (\url{https://etymo.io}), a visual search engine for data scientists.

\section{Time-Preserving walks}
\label{sec:time-pres-walks}

For example, in Figure \ref{fig:eg_shortest_walk}
$\langle (1, t_1), (2, t_1) \rangle$ is a static edge at time stamp $t_1$ and $\langle (3, t_2), (3, t_3) \rangle$ is a causal edge between time stamp $t_2$ and time stamp $t_3$.
The difference between static edges and causal edges were first explicitly considered in \cite{chen16}.

\begin{figure}[h]
 \begin{center}
    \begin{tikzpicture}[scale=.38, line width =1.5pt]
  \node[circle,fill=green1, minimum size=0.2cm] (n7) at (-5,7) {2};
  \node[circle,draw=red1, minimum size=0.2cm] (n8) at (-10,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n10) at (-8,10) {1};
  \node[circle,draw=red1, minimum size=0.2cm] (n6) at (3,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n4) at (-2,5) {3};
  \node[circle,fill=green1, minimum size=0.2cm] (n1) at (0,10) {1};
   \node[circle,fill=green1, minimum size=0.2cm] (n11) at (11,7) {2};
  \node[circle,fill=green1, minimum size=0.2cm] (n12) at (6,5)  {3};
  \node[circle,draw=red1, minimum size=0.2cm] (n14) at (8,10) {1};
  \node (w1) at (-6, 9) {$w_1$};
  \node (w2) at (3, 13) {$w_2$};
  \node (w3) at (-2, 8) {$w_3$};
  \node (w4) at (2, 4) {$w_4$};
  \node (w5) at (8.5, 7) {$w_5$};s
  \foreach \from/\to in {n10/n7, n1/n4, n11/n12}
   \draw[every edge,->] (\from) -- (\to);
 \draw[dotted,->](n4) -- (n12);
\draw[dotted,->](n7) to[out=80, in=40](n11);
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(4,3) -- (-2,3) node [midway,yshift=-0.6cm]{ $t_2$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(-4,3) -- (-10.5,3) node [midway,yshift=-0.6cm]{ $t_1$};
\draw [decorate,decoration={brace,amplitude=5pt},xshift=-4pt,yshift=0pt]
(11.5,3) -- (5.5,3) node [midway,yshift=-0.6cm]{ $t_3$};
    \end{tikzpicture}
\end{center}
\caption{An evolving directed graph with 3 time stamps $t_1$, $t_2$ and $t_3$.
At each time stamp, the evolving graph is represented as a graph.
The green filled circles represent active nodes while the red circles represent
inactive nodes. Directed edges in each time stamp are shown as black arrows and directed edges between graphs are shown as dotted arrows, where $w_1, w_2 \ldots w_5$ are edge weights.}
\label{fig:eg_shortest_walk}
\end{figure}

It is natural to define the length of a walk on a static graph as the number of nodes travelled through the walk. For the weighted graph case, the length of a walk is the total weight of all the edges in the walk. We observe that static edges link nodes in space while causal edges link nodes in time. Hence we would like to distinguish temporal distance from spatial distance on an evolving graph: the length of a static edge is equal to the edge weight (or $1$ in the unweighted graph case); the length of a causal edge is the time distance between the two nodes. In the simplest case, the time-respecting walk
$\langle (1, t_1) ,(2, t_1) , (2, t_3), (3, t_3)\rangle$ in Figure \ref{fig:eg_shortest_walk} has length $4$ in total because two static edges $\langle (1, t_1) (2, t_1) \rangle$ and $\langle (2, t_3), (3, t_3) \rangle$ have length one, whereas $\langle (2, t_1) (2, t_3) \rangle$ traverse $2$ time stamps so has length $2$.

We now describe time-preserving walks in more detail. We follow the definition of
\emph{evolving graph}, \emph{temporal node} and \emph{active node} in \cite{chen16}.

@@

Note the node sets $V^{[t]}$ can change over time, i.e., nodes may appear or disappear at a particular time stamp.
For example, in Figure \ref{fig:eg_shortest_walk}, at time stamp $t_1$, $V^{[1]} = \langle 1, 2 \rangle$ and $E^{[1]} = \langle (1,2) \rangle$. Each graph $G^{[t]}$ can be represented by its adjacency matrix $A^{[t]}$.
We can represent $G_n$ by a list of adjacency matrices $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{[n]} \rangle$.

@@

In Figure \ref{fig:eg_shortest_walk}, the filled green circles are active nodes. For the adjacency matrix representation, a node $i$ is active at time stamp $t$, then
the $i$th row or the $i$th column of $A^{[t]}$ has at least one none-zero entry.
 In general, a time-preserving walk can be defined as follows.

\begin{definition}
A \textbf{time-preserving walk}  of length $m$ on an evolving graph $G_n$ is a time-ordered sequence of active nodes, $\langle (v_1, t_1), (v_2, t_2), \ldots, (v_m, t_m) \rangle$, where $t_1 \le t_2 \le \cdots \le t_m$ and
$v_i = v_j$ if and only if $t_i \ne t_j$.
\end{definition}

@@

We can denote a \emph{weighted static edge} as $\langle (v_i, t), (v_j, t), w_s \rangle$ and a \emph{weighted causal edge} as $\langle (v, t_i), (v, t_j), w_t \rangle$.
Here $w_s$ represents the spatial distance between the two nodes $v_i$ and $v_j$; $w_t$ represents the
temporal distance of $v$ at different time stamps.
In Figure \ref{fig:eg_shortest_walk}, the black arrows represent static edges and the dotted arrows represent causal edges.
Suppose we assign all the static edge weight to be $1$ and
assign the causal edge to be the number of time stamps between the pair. Then in Figure \ref{fig:eg_shortest_walk}
the edge weight between $(1, t_1)$ and $(2, t_1)$ is $1$ and the edge weight between $(2, t_1)$ and $(2, t_3)$
 is $2$.

A time-preserving walk can traverse both causal edges and static edges.
We will see in the next section that evolving centrality measures make implicit assumptions about the type of edges in the
time-preserving walk that is used.

\section{Centrality Measures}
\label{sec:topol-temp-flow}

In \cite{tang10s}, a temporal shortest walk is measured in time units only, i.e., if node
$v_i$ can reach node $v_j$ in $k$ time steps then the distance between node
$v_i$ and $v_j$ is $k$. We can apply temporal breadth first
search (BFS) \cite{chen16} to discover the temporal walks and then only take account of the temporal difference between the first node $(v_1, t_1)$ and the last node $(v_n, t_n)$, i.e., $t_n - t_1$.

If we use this definition of temporal walk for the temporal Katz centrality introduced in
\cite{grindrod11}, we have a new centrality measure.
In the matrix form, the nonzero entries of $A^{[t_i]} A^{[t_{i+1}]}$ counts all the temporal walks of length $1$, and
$A^{[t_i]}\cdots A^{[t_j]}$, where $i < j$ counts all the temporal walks of length $j -i$.
For example, $A^{[1]}A^{[4]}$, $A^{[1]}A^{[1]}A^{[4]}$ and $A^{[1]}A^{[2]}A^{[3]}A^{[4]}$ are all walks of length 3.
In the original Katz centrality, the `attenuation' factor $\alpha$ is imposed on the
number of walks, i.e., longer walks has smaller effect to the overall rating.
Hence the final result is the summation of all products of the form
\[
\alpha^k A^{[i]} \cdots A^{[i+k]}
\]
% Note this can not be written in \eqref{eq:katz}, as discussed in Section \ref{sec:temp-katz-centr}.

Another definition of temporal walk was introduced by \cite{chen16}. Here
both static edges and causal edges are considered. We can represent an evolving graph by a block adjacency matrix, as shown in Section \ref{sec:centr-block-adjac}. Centrality on this block matrix gives the importance of each node at all time stamps. We can average these rating at different time stamps to
get the final rating.

We review a few well-known measures of evolving graph centrality.
(Design a simple example and discuss it in each case below.)

\subsection{Temporal Katz Centrality and Communicability Betweenness Centrality}
\label{sec:temp-katz-centr}

@@

\subsection{Temporal Betweenness and Closeness Centrality}
\label{sec:temp-betw-centr}

Recall that betweenness centrality of a node in a static graph is defined by counting the fraction of shortest walks that traverse it.
We could extend
betweenness centrality on evolving graphs by replacing shortest walks with temporal shortest walks \cite{nicosia13}.
\begin{equation}
  \label{eq:bet1}
C_i^{betweenness} = \sum_{j \in V}\sum_{k \in V, k \ne j}\frac{\alpha_{jk}(i)}{\alpha_{jk}},
\end{equation}
where $\alpha_{jk}$ represents the number of temporal shortest walks from node $j$ to node $k$ and $\alpha_{jk}(i)$ represents the number of temporal shortest walks from node $j$ to node $k$ that pass through the node $i$.

One may want to take into account the length of time a node passes a message. The temporal betweenness centrality of the node $i$ at time stamp $t_m$ is defined as
\begin{equation}
  \label{eq:bet2}
  C_i^{betweenness}(t_m) = \sum_{j\ne i}\sum_{k\ne j, k\ne i}\frac{U_{j,k,t_m}(i)}{\alpha_{j,k}},
\end{equation}
where, as before, $\alpha_{j,k}$ represents the number of temporal shortest walks from $j$ to $k$ and $U_{j,k,t_m}(i)$ represents the number of temporal shortest walks from $j$ to $k$ in which node $i$ is traversed from
the walk in the time stamp $t_m$ or a previous time stamp $t_{m-1}$. Then
average temporal betweenness of node $i$ is then defined as
$$
  C_i^{betweenness} = \frac{1}{M}\sum_m C_i^{betweenness}(t_m).
$$
Note \eqref{eq:bet1} only takes account of static edges while \eqref{eq:bet2} takes account of both static edges and causal edges.
